---
title: "Assignment 1"
author: "Michal Dzwolak"
date: "29 June 2017"
output: html_document
---


###What is mean total number of steps taken per day?

Table with sum, mean and median of steps taken per day.

```{r pressure, echo=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(knitr)

up <- read.csv('activity.csv')

up1 <- up %>% select(steps, date) %>% group_by(date) %>% summarise(Mean=round(mean(steps),2), Sum=sum(steps), Median=median(steps))

pl <- ggplot(up, aes(x=up$date, y=up$steps)) + geom_bar(stat = 'identity') + 
  theme(axis.text.x=element_text(angle=90, hjust=1)) + ylab('Sum of steps taken') + xlab('Date')

kable(up1, format = 'html')

print(pl)

```



###What is the average daily activity pattern?

```{r firstplot, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)

up <- na.omit(read.csv('activity.csv'))

up1 <- up %>% group_by(interval) %>% summarise(Avg=round(mean(steps),2))

up2 <- up %>% group_by(interval) %>% summarise(Sum=sum(steps)) %>% filter(Sum==max(Sum))

inter <- up2$interval

suma <- up2$Sum

pl <- plot(up1$interval, up1$Avg, type = 'l', xlab = '5-minute intervwal', ylab = 'Avg. of steps')

```

The maximum number of steps we can wind in `r inter` interval and its equal to `r suma` steps. 


###Imputing missing values
```{r secondplt, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(knitr)

up <- read.csv('activity.csv')
up2 <-na.omit(read.csv('activity.csv'))

countna <- up %>% select(steps) %>% filter(is.na(steps))

dataset <- up %>% select(steps, date, interval) %>% filter(!is.na(steps)) %>% group_by(interval) %>% summarise(Avg=round(mean(steps),2))

nareplacement <- up %>% select(steps, date, interval) %>% filter(is.na(steps))

nareplacement <- left_join(nareplacement, dataset, by = "interval") %>% select(Avg, date, interval) 

nareplacement <- rename(nareplacement, steps = Avg)

nareplacement <- rbind(up2,nareplacement)

pl <- ggplot(up, aes(x=nareplacement$date, y=nareplacement$steps)) + geom_bar(stat = 'identity') + 
  theme(axis.text.x=element_text(angle=90, hjust=1)) + ylab('Sum of steps taken') + xlab('Date')

datatotable <- nareplacement %>% group_by(date) %>% summarise(Avg=round(mean(steps),2), Median=round(median(steps),2))

kable(datatotable, format = "html")

print(pl)
```

 - We can see that days with NA results from the previous graph now have results. These results are averages of each interval.  



###Imputing missing values

```{r thirdplot, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(knitr)

up <- read.csv('activity.csv')
up2 <-na.omit(read.csv('activity.csv'))

countna <- up %>% select(steps) %>% filter(is.na(steps))

dataset <- up %>% select(steps, date, interval) %>% filter(!is.na(steps)) %>% group_by(interval) %>% summarise(Avg=round(mean(steps),2))

nareplacement <- up %>% select(steps, date, interval) %>% filter(is.na(steps))

nareplacement <- left_join(nareplacement, dataset, by = "interval") %>% select(Avg, date, interval) 

nareplacement <- rename(nareplacement, steps = Avg)

narep <- rbind(up2,nareplacement)

narep <- narep %>% mutate(weekday=weekdays(as.Date(date))) %>% 
  mutate(split=ifelse(weekday=='Sunday' | weekday=='Saturday','Weekend','Workday'))

narep <- narep %>% select(steps, interval, split) %>% group_by(interval, split) %>% summarise(Avg=round(mean(steps),2))

pl <- ggplot(narep, aes(x=interval, y=Avg)) + geom_line() + facet_grid(split~.) + theme_bw()  

print(pl)

```